|

### 2021-03-21 documentation/paper.Rmd
Fixed issue with Cabin font and minor formatting issue in README
Moved some functions to an internal folder to remove cumbersome dependencies like Stan
Various fixes

pbp_long <- 
  pollbasepro %>% 
  pivot_longer(
    cols = c(-date, -election, -govt, -week, -month, -quarter, -year, -con_ldr, -lab_ldr, -lib_ldr),
    names_to = c("party", ".value"),
    names_sep = "_",
  ) %>%
  mutate(
    party =
      case_when(
        party == "con" ~ "Conservative Party",
        party == "lab" ~ "Labour Party",
        party == "lib" ~ "Liberals (Various Forms)"
      ) %>% 
      factor(
        levels = 
          c("Conservative Party",
            "Labour Party",
            "Liberals (Various Forms)"
          )
      )
  )

@@

timeline <-
  load_timeline() %>%
  select(
    date = polldate,
    elecdate,
    country,
    party = partyid,
    vote = poll_
  ) %>%
  filter(country == "United Kingdom") %>% 
  na.omit()

@@

death_mod <- readRDS(here("models", paste0("death_mod_", packageVersion("britpol"), ".rds")))

# Load correlation models

cor_all <- readRDS(here("models", paste0("cor_all_", packageVersion("britpol"), ".rds")))
cor_con <- readRDS(here("models", paste0("cor_con_", packageVersion("britpol"), ".rds")))
cor_lab <- readRDS(here("models", paste0("cor_lab_", packageVersion("britpol"), ".rds")))
cor_lib <- readRDS(here("models", paste0("cor_lib_", packageVersion("britpol"), ".rds")))

@@

tl <-
  load_timeline() %>%
  select(
    date = polldate,
    elecdate,
    country,
    party = partyid,
    vote = poll_
  ) %>%
  filter(country == "United Kingdom") %>%
  mutate(
    vote = vote/100,
    party = case_when(party == 1 ~ "con", party == 2 ~ "lab", party == 3 ~ "lib"),
    polldate = as_date(date)
  ) %>%
  na.omit()

@@

ggplot() +
  geom_density_2d(
    data = con %>% mutate(facet = "Conservative"),
    mapping = aes(x = vote, y = con_est),
    colour = pty_cols[1],
    alpha = .4
  ) +
  stat_smooth(
    data = con %>% mutate(facet = "Conservative"),
    mapping = aes(x = vote, y = con_est),
    colour = "black",
    fill = pty_cols[1],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = con %>% mutate(facet = "Conservative"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           in_text(pluck(posterior_samples(cor_con, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_con$data$con_est, cor_con$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_con$data$con_est, cor_con$data$vote), 2)
  #           )
  #       ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  geom_density_2d(
    data = lab %>% mutate(facet = "Labour"),
    mapping = aes(x = vote, y = lab_est),
    colour = pty_cols[2],
    alpha = .4
  ) +
  stat_smooth(
    data = lab %>% mutate(facet = "Labour"),
    mapping = aes(x = vote, y = lab_est),
    colour = "black",
    fill = pty_cols[2],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = lab %>% mutate(facet = "Labour"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           in_text(pluck(posterior_samples(cor_lab, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_lab$data$lab_est, cor_lab$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_lab$data$lab_est, cor_lab$data$vote), 2)
  #           )
  #     ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  geom_density_2d(
    data = lib %>% mutate(facet = "Liberal (Various Forms)"),
    mapping = aes(x = vote, y = lib_est),
    colour = pty_cols[3],
    alpha = .4
  ) +
  stat_smooth(
    data = lib %>% mutate(facet = "Liberal (Various Forms)"),
    mapping = aes(x = vote, y = lib_est),
    colour = "black",
    fill = pty_cols[3],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = lib %>% mutate(facet = "Liberal (Various Forms)"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           in_text(pluck(posterior_samples(cor_lib, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_lib$data$lib_est, cor_lib$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_lib$data$lib_est, cor_lib$data$vote), 2)
  #           )
  #     ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  facet_wrap(~ facet) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Jennings and Wlezien's 'Timeline of Elections' Data", y = "PollBasePro Estimates") +
  theme_bailey()
```

@@

We validate them against Jennings and Wlezien's "Timeline of Elections" dataset [-@jennings2016a]. These data contain `r format(nrow(timeline[timeline$party == 1, ]), big.mark = ",")` polls from Britain from `r str_remove(format(min(as_date(timeline$date)), "%d %B %Y"), "^0")` to `r str_remove(format(max(as_date(timeline$date)), "%d %B %Y"), "^0")`. Given that our data are so comprehensive, it is likely that most polls appear in both datasets. Still, the Timeline data provide a good test as Jennings and Wlezien compiled them independently. As figure \@ref(fig:val-plot) shows, our estimates are well validated. Correlations between the two series are strong and positive. Their mean absolute error (MAE) and root-mean-square error (RMSE) are also low in all cases. The Conservatives showed a correlation of `r in_text(pluck(posterior_samples(cor_con, "rescor"), 1)*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(mae(cor_con$data$con_est*100, cor_con$data$vote*100), 2), " percentage points")`, and an RMSE of `r round(rmse(cor_con$data$con_est, cor_con$data$vote), 2)`; Labour, a correlation of `r in_text(pluck(posterior_samples(cor_lab, "rescor"), 1)*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(mae(cor_lab$data$lab_est*100, cor_lab$data$vote*100), 2), " points")`, and an RMSE of `r round(rmse(cor_lab$data$lab_est, cor_lab$data$vote), 2)`; and the Liberals, a correlation of `r in_text(pluck(posterior_samples(cor_lib, "rescor"), 1)*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(mae(cor_lib$data$lib_est*100, cor_lib$data$vote*100), 2), " points")`, and an RMSE of `r round(rmse(cor_lib$data$lib_est, cor_lib$data$vote), 2)`.

@@

As well as summarising facts about the parties, our estimates also allow us to make general claims about the relationships that exist *between* the parties. As figure \@ref(fig:cor-plot) shows, these are all negative. Consider the effect of rising Liberal support on the Conservatives and Labour. The correlation between the Liberals and the Conservatives is `r in_text(pluck(posterior_samples(cor_all, "rescor__libest__conest"), 1)*100, suffix = "%", inside = F)` and between the Liberals and Labour `r in_text(pluck(posterior_samples(cor_all, "rescor__libest__labest"), 1)*100, suffix = "%", inside = F)`. Contrary to popular arguments that the Liberals serve to split the centre-left vote [e.g. @jenkins2019], our figures suggest that rising Liberal support has tended to hurt the Conservatives more than Labour. Interestingly, the correlation between Labour and the Conservatives is also the smallest, at `r in_text(pluck(posterior_samples(cor_all, "rescor__conest__labest"), 1)*100, suffix = "%", inside = F)`. This is perhaps unsurprising, given that the two parties represent different sides of the left-right spectrum. Even so, that both parties have tended to target the other's vote still makes strategic sense: each party's pool of potential voters is almost always larger than the pool of potential Liberal voters.

@@

We can also use our estimates to make specific claims about British politics. For example, we can assert with a reasonable degree of certainty that the `r pbp_long$party[which.max(pbp_long$est)]` received the largest degree of support of any party in Britain between `r format(min(pollbasepro$date), "%Y")` and `r format(max(pollbasepro$date), "%Y")` on `r str_remove(format(pbp_long$date[which.max(pbp_long$est)], "%d %B %Y"), "^0")` when our estimates show that `r scales::percent(pbp_long$est[which.max(pbp_long$est)], accuracy = 0.1)` (95% CI: `r paste(scales::percent(pbp_long$est[which.max(pbp_long$est)] - qnorm(0.975)*pbp_long$err[which.max(pbp_long$est)], accuracy = 0.1), "to", scales::percent(pbp_long$est[which.max(pbp_long$est)] + qnorm(0.975)*pbp_long$err[which.max(pbp_long$est)], accuracy = 0.1))`) of the electorate would vote for them at the next election. Similarly, we can assert that Labour's peak in the polls came on `r str_remove(format(pollbasepro$date[which.max(pollbasepro$lab_est)], "%d %B %Y"), "^0")` when `r scales::percent(max(pollbasepro$lab_est), accuracy = .1)` (95% CI: `r paste(scales::percent(max(pollbasepro$lab_est) - qnorm(0.975)*pollbasepro$lab_err[which.max(pollbasepro$lab_est)], accuracy = 0.1), "to", scales::percent(max(pollbasepro$lab_est) + qnorm(0.975)*pollbasepro$lab_err[which.max(pollbasepro$lab_est)], accuracy = 0.1))`) of the electorate intended to back them.

These data and their summaries raise an interesting question: if Labour and the Conservatives have tended to be so well-matched in the polls, why have the Conservatives done so much better at election time? Despite leading in the polls `r str_remove(lead$Leader[2], " of days")` of the time, Labour has gained the highest share of the vote in only `r pbp_long %>% filter(err == 0) %>% group_by(election) %>% summarise(winner = party[which.max(est)], .groups = "drop") %>% summarise(lab = length(winner[winner == "Labour Party"]), .groups = "drop") %>% pluck(1)` of the period's `r pbp_long %>% filter(err == 0) %>% group_by(election) %>% summarise(winner = party[which.max(est)], .groups = "drop") %>% nrow()` general elections (`r scales::percent((pbp_long %>% filter(err == 0) %>% group_by(election) %>% summarise(winner = party[which.max(est)], .groups = "drop") %>% summarise(lab = length(winner[winner == "Labour Party"]), .groups = "drop") %>% pluck(1))/(pbp_long %>% filter(err == 0) %>% group_by(election) %>% summarise(winner = party[which.max(est)], .groups = "drop") %>% nrow()))`). This phenomenon is not entirely unprecedented. @jackman1994 shows that the Australian Labor Party has suffered at elections due to pervasive electoral bias. Though this might also be the case in Britain, our figures reflect *vote* shares, not *seat* shares. As such, electoral bias should be a concern only insofar as it affects the parties' overall popularity. Other factors likely explain Labour's poor performance: Britain's press leans right and tends to support the Conservatives; successive Conservative governments could have scheduled elections to maximise their chances; partisan non-response could have led Conservative voters to drop out of polls in the period between elections; or, most simply, the Conservatives might just be better at campaigning.

@@

plot1 <- 
  pollbasepro %>% 
  ggplot(
    aes(
      x = lib_est,
      y = con_est
    )
  ) +
  geom_density_2d(
    colour = bailey_colours("grey4"),
    alpha = .4
    ) +
  stat_smooth(
    method = "lm",
    formula = "y ~ x",
    colour = bailey_colours("black"),
    se = TRUE
  ) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Liberal Support", y = "Conservative Support") +
  annotate(
    "text",
    label = 
      in_text(pluck(posterior_samples(cor_all, "rescor__libest__conest"), 1)*100, suffix = "%", inside = F),
    x = .6, y = .1,
    hjust = 1,
    size = 2,
    family = "Cabin"
    ) +
  theme_bailey()

@@

plot2 <- 
  pollbasepro %>% 
  ggplot(
    aes(
      x = lib_est,
      y = lab_est
    )
  ) +
  geom_density_2d(
    colour = bailey_colours("grey4"),
    alpha = .4
    ) +
  stat_smooth(
    method = "lm",
    formula = "y ~ x",
    colour = bailey_colours("black"),
    se = TRUE
  ) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Liberal Support", y = "Labour Support") +
  annotate(
    "text",
    label = 
      in_text(pluck(posterior_samples(cor_all, "rescor__libest__labest"), 1)*100, suffix = "%", inside = F),
    x = .6, y = .1,
    hjust = 1,
    size = 2,
    family = "Cabin"
    ) +
  theme_bailey()

@@

plot3 <- 
  pollbasepro %>% 
  ggplot(
    aes(
      x = lab_est,
      y = con_est
    )
  ) +
  geom_density_2d(
    colour = bailey_colours("grey4"),
    alpha = .4
    ) +
  stat_smooth(
    method = "lm",
    formula = "y ~ x",
    colour = bailey_colours("black"),
    se = TRUE
  ) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Labour Support", y = "Conservative Support") +
  annotate(
    "text",
    label = 
    in_text(pluck(posterior_samples(cor_all, "rescor__conest__labest"), 1)*100, suffix = "%", inside = F),
    x = .6, y = .1,
    hjust = 1,
    size = 2,
    family = "Cabin"
    ) +
  theme_bailey()

@@

Figure \@ref(fig:death-plot) shows that the effect is more "rally 'round the flag" than "reward and punishment". That said, our model suggests at least some preliminary evidence of a transition from one mechanism to the other. In March 2020, when the pandemic started to gain pace in Britain, there was a strong positive relationship between change in the number of deaths attributable to COVID-19 and change in support for the governing Conservative Party. For every 100 extra COVID-related deaths over the past 4 weeks, Conservative support over the same period increased on average by `r in_text((pluck(posterior_samples(death_mod, "b_deaths"), 1) + pluck(posterior_samples(death_mod, "b_time:deaths"), 1)*2)*100, inside = F, text = " points")`. Six months later, in September 2020, this relationship had weakened markedly. Though still positive, it had fallen to only `r in_text((pluck(posterior_samples(death_mod, "b_deaths"), 1) + pluck(posterior_samples(death_mod, "b_time:deaths"), 1)*8)*100, inside = F, text = " points")`. A year after the pandemic began, the relationship had reversed. In March 2021, every 100 additional COVID-related deaths over the past 4 weeks led Conservative support to fall on average by `r in_text((pluck(posterior_samples(death_mod, "b_deaths"), 1) + pluck(posterior_samples(death_mod, "b_time:deaths"), 1)*14)*100, inside = F, text = " points")`.

@@

cond_fx %>% 
  ggplot(
    aes(
      x = deaths,
      y = estimate__,
      ymin = estimate__ - qnorm(.975)*se__,
      ymax = estimate__ + qnorm(.975)*se__
    )
  ) +
  facet_wrap(~ cond__) +
  geom_vline(
    xintercept = 0,
    colour = bailey_colours("grey4"),
    linetype = "dotted"
  ) +
  geom_ribbon(
    colour = NA,
    fill = bailey_colours("red"),
    alpha = .2) +
  geom_line(colour = bailey_colours("red")) +
  scale_y_continuous(
    breaks = seq(-.1, .1, by = .05),
    labels = scales::percent_format(suffix = "pts", accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(-10, 10, by = 5),
    labels = scales::number_format(scale = 100)
  ) +
  coord_cartesian(
    ylim = c(-.1, .1),
    xlim = c(-10, 10)
  ) +
  labs(
    y = "Change in Con. Support, Past 4 Weeks",
    x = "Change in Deaths Attributable to COVID-19, Past 4 Weeks"
  ) +
  theme_bailey()
```

@@

save_info(path = here("sessions", "006_paper.txt"))

>>>>

pbp_long <- 
  pollbasepro %>% 
  pivot_longer(
    cols = -date,
    names_to = c("party", ".value"),
    names_sep = "_",
  ) %>%
  mutate(
    party =
      case_when(
        party == "con" ~ "Conservative Party",
        party == "lab" ~ "Labour Party",
        party == "lib" ~ "Liberals (Various Forms)"
      ) %>% 
      factor(
        levels = 
          c("Conservative Party",
            "Labour Party",
            "Liberals (Various Forms)"
          )
      )
  )

@@

timeline <-
  britpol:::load_timeline() %>%
  select(
    date = polldate,
    elecdate,
    country,
    party = partyid,
    vote = poll_
  ) %>%
  filter(country == "United Kingdom") %>% 
  na.omit()

@@

death_mod <- 
  readRDS(
    here(
        "documentation",
        "_assets",
        paste0("death_mod_", packageVersion("britpol"), ".rds")
      )
  )

@@

tl <-
  britpol:::load_timeline() %>%
  select(
    date = polldate,
    elecdate,
    country,
    party = partyid,
    vote = poll_
  ) %>%
  filter(country == "United Kingdom") %>%
  mutate(
    vote = vote/100,
    party = case_when(party == 1 ~ "con", party == 2 ~ "lab", party == 3 ~ "lib"),
    polldate = as_date(date)
  ) %>%
  na.omit()

@@

ggplot() +
  geom_density_2d(
    data = con %>% mutate(facet = "Conservative"),
    mapping = aes(x = vote, y = con_est),
    colour = pty_cols[1],
    alpha = .4
  ) +
  stat_smooth(
    data = con %>% mutate(facet = "Conservative"),
    mapping = aes(x = vote, y = con_est),
    colour = "black",
    fill = pty_cols[1],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = con %>% mutate(facet = "Conservative"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_con, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_con$data$con_est, cor_con$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_con$data$con_est, cor_con$data$vote), 2)
  #           )
  #       ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  geom_density_2d(
    data = lab %>% mutate(facet = "Labour"),
    mapping = aes(x = vote, y = lab_est),
    colour = pty_cols[2],
    alpha = .4
  ) +
  stat_smooth(
    data = lab %>% mutate(facet = "Labour"),
    mapping = aes(x = vote, y = lab_est),
    colour = "black",
    fill = pty_cols[2],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = lab %>% mutate(facet = "Labour"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_lab, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_lab$data$lab_est, cor_lab$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_lab$data$lab_est, cor_lab$data$vote), 2)
  #           )
  #     ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  geom_density_2d(
    data = lib %>% mutate(facet = "Liberal (Various Forms)"),
    mapping = aes(x = vote, y = lib_est),
    colour = pty_cols[3],
    alpha = .4
  ) +
  stat_smooth(
    data = lib %>% mutate(facet = "Liberal (Various Forms)"),
    mapping = aes(x = vote, y = lib_est),
    colour = "black",
    fill = pty_cols[3],
    method = "lm",
    formula = y ~ x
  ) +
  # geom_text(
  #   data = lib %>% mutate(facet = "Liberal (Various Forms)"),
  #   mapping = 
  #     aes(
  #       x = .6,
  #       y = .05,
  #       label = 
  #         paste0(
  #           britpol:::in_text(pluck(posterior_samples(cor_lib, "rescor"), 1)*100, suffix = "%", inside = F),
  #           "\n",
  #           "MAE = ", round(mae(cor_lib$data$lib_est, cor_lib$data$vote), 2),
  #           ", RMSE = ", round(rmse(cor_lib$data$lib_est, cor_lib$data$vote), 2)
  #           )
  #     ),
  #   hjust = 1,
  #   size = 2,
  #   family = "Cabin Regular"
  # ) +
  facet_wrap(~ facet) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Jennings and Wlezien's 'Timeline of Elections' Data", y = "PollBasePro Estimates") +
  theme_bailey()
```

@@

We validate them against Jennings and Wlezien's "Timeline of Elections" dataset [-@jennings2016a]. These data contain `r format(nrow(timeline[timeline$party == 1, ]), big.mark = ",")` polls from Britain from `r str_remove(format(min(as_date(timeline$date)), "%d %B %Y"), "^0")` to `r str_remove(format(max(as_date(timeline$date)), "%d %B %Y"), "^0")`. Given that our data are so comprehensive, it is likely that most polls appear in both datasets. Still, the Timeline data provide a good test as Jennings and Wlezien compiled them independently. As figure \@ref(fig:val-plot) shows, our estimates are well validated. Correlations between the two series are strong and positive. Their mean absolute error (MAE) and root-mean-square error (RMSE) are also low in all cases. The Conservatives showed a correlation of `r britpol:::in_text(cor_mods$tl_con*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_con, 2), " percentage points")`, and an RMSE of `r round(cor_mods$rmse_con, 2)`; Labour, a correlation of `r britpol:::in_text(cor_mods$tl_lab*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_lab, 2), " points")`, and an RMSE of `r round(cor_mods$rmse_lab, 2)`; and the Liberals, a correlation of `r britpol:::in_text(cor_mods$tl_lib*100, suffix = "%", inside = F)`, an MAE of `r paste0(round(cor_mods$mae_lib, 2), " points")`, and an RMSE of `r round(cor_mods$rmse_lib, 2)`.

@@

As well as summarising facts about the parties, our estimates also allow us to make general claims about the relationships that exist *between* the parties. As figure \@ref(fig:cor-plot) shows, these are all negative. Consider the effect of rising Liberal support on the Conservatives and Labour. The correlation between the Liberals and the Conservatives is `r britpol:::in_text(cor_mods$lib_con*100, suffix = "%", inside = F)` and between the Liberals and Labour `r britpol:::in_text(cor_mods$lib_lab*100, suffix = "%", inside = F)`. Contrary to popular arguments that the Liberals serve to split the centre-left vote [e.g. @jenkins2019], our figures suggest that rising Liberal support has tended to hurt the Conservatives more than Labour. Interestingly, the correlation between Labour and the Conservatives is also the smallest, at `r britpol:::in_text(cor_mods$con_lab*100, suffix = "%", inside = F)`. This is perhaps unsurprising, given that the two parties represent different sides of the left-right spectrum. Even so, that both parties have tended to target the other's vote still makes strategic sense: each party's pool of potential voters is almost always larger than the pool of potential Liberal voters.

@@

We can also use our estimates to make specific claims about British politics. For example, we can assert with a reasonable degree of certainty that the `r pbp_long$party[which.max(pbp_long$est)]` received the largest degree of support of any party in Britain between `r format(min(pollbasepro$date), "%Y")` and `r format(max(pollbasepro$date), "%Y")` on `r str_remove(format(pbp_long$date[which.max(pbp_long$est)], "%d %B %Y"), "^0")` when our estimates show that `r scales::percent(pbp_long$est[which.max(pbp_long$est)], accuracy = 0.1)` (95% CI: `r paste(scales::percent(pbp_long$est[which.max(pbp_long$est)] - qnorm(0.975)*pbp_long$err[which.max(pbp_long$est)], accuracy = 0.1), "to", scales::percent(pbp_long$est[which.max(pbp_long$est)] + qnorm(0.975)*pbp_long$err[which.max(pbp_long$est)], accuracy = 0.1))`) of the electorate would vote for them at the next election. Similarly, we can assert that the Conservative's peak in the polls came on `r str_remove(format(pollbasepro$date[which.max(pollbasepro$con_est)], "%d %B %Y"), "^0")` when `r scales::percent(max(pollbasepro$con_est), accuracy = .1)` (95% CI: `r paste(scales::percent(max(pollbasepro$con_est) - qnorm(0.975)*pollbasepro$con_err[which.max(pollbasepro$con_est)], accuracy = 0.1), "to", scales::percent(max(pollbasepro$con_est) + qnorm(0.975)*pollbasepro$con_err[which.max(pollbasepro$con_est)], accuracy = 0.1))`) of the electorate intended to back them.

These data and their summaries raise an interesting question: if Labour and the Conservatives have tended to be so well-matched in the polls, why have the Conservatives done so much better at election time? Despite leading in the polls `r str_remove(lead$Leader[2], " of days")` of the time, Labour has gained the highest share of the vote in only `r pbp_long %>% filter(err == 0) %>% add_elections() %>% group_by(last_elec) %>% summarise(winner = party[which.max(est)], .groups = "drop") %>% summarise(lab = length(winner[winner == "Labour Party"]), .groups = "drop") %>% pluck(1)` of the period's `r pbp_long %>% filter(err == 0) %>% add_elections() %>%  group_by(last_elec) %>% summarise(winner = party[which.max(est)], .groups = "drop") %>% nrow()` general elections (`r scales::percent((pbp_long %>% filter(err == 0) %>% add_elections() %>% group_by(last_elec) %>% summarise(winner = party[which.max(est)], .groups = "drop") %>% summarise(lab = length(winner[winner == "Labour Party"]), .groups = "drop") %>% pluck(1))/(pbp_long %>% filter(err == 0) %>% add_elections() %>% group_by(last_elec) %>% summarise(winner = party[which.max(est)], .groups = "drop") %>% nrow()))`). This phenomenon is not entirely unprecedented. @jackman1994 shows that the Australian Labor Party has suffered at elections due to pervasive electoral bias. Though this might also be the case in Britain, our figures reflect *vote* shares, not *seat* shares. As such, electoral bias should be a concern only insofar as it affects the parties' overall popularity. Other factors likely explain Labour's poor performance: Britain's press leans right and tends to support the Conservatives; successive Conservative governments could have scheduled elections to maximise their chances; partisan non-response could have led Conservative voters to drop out of polls in the period between elections; or, most simply, the Conservatives might just be better at campaigning.

@@

plot1 <- 
  pollbasepro %>% 
  ggplot(
    aes(
      x = lib_est,
      y = con_est
    )
  ) +
  geom_density_2d(
    colour = britpol:::bailey_colours("grey4"),
    alpha = .4
    ) +
  stat_smooth(
    method = "lm",
    formula = "y ~ x",
    colour = britpol:::bailey_colours("black"),
    se = TRUE
  ) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Liberal Support", y = "Conservative Support") +
  annotate(
    "text",
    label = britpol:::in_text(cor_mods$lib_con*100, suffix = "%", inside = F),
    x = .6, y = .1,
    hjust = 1,
    size = 2,
    family = "Cabin"
    ) +
  theme_bailey()

@@

plot2 <- 
  pollbasepro %>% 
  ggplot(
    aes(
      x = lib_est,
      y = lab_est
    )
  ) +
  geom_density_2d(
    colour = britpol:::bailey_colours("grey4"),
    alpha = .4
    ) +
  stat_smooth(
    method = "lm",
    formula = "y ~ x",
    colour = britpol:::bailey_colours("black"),
    se = TRUE
  ) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Liberal Support", y = "Labour Support") +
  annotate(
    "text",
    label = britpol:::in_text(cor_mods$lib_lab*100, suffix = "%", inside = F),
    x = .6, y = .1,
    hjust = 1,
    size = 2,
    family = "Cabin"
    ) +
  theme_bailey()

@@

plot3 <- 
  pollbasepro %>% 
  ggplot(
    aes(
      x = lab_est,
      y = con_est
    )
  ) +
  geom_density_2d(
    colour = britpol:::bailey_colours("grey4"),
    alpha = .4
    ) +
  stat_smooth(
    method = "lm",
    formula = "y ~ x",
    colour = britpol:::bailey_colours("black"),
    se = TRUE
  ) +
  scale_y_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(0, .6, by = .2),
    labels = scales::percent_format(accuracy = 1)
  ) +
  coord_cartesian(
    ylim = c(0, 0.62),
    xlim = c(0, 0.62)
  ) +
  labs(x = "Labour Support", y = "Conservative Support") +
  annotate(
    "text",
    label = britpol:::in_text(cor_mods$con_lab*100, suffix = "%", inside = F),
    x = .6, y = .1,
    hjust = 1,
    size = 2,
    family = "Cabin"
    ) +
  theme_bailey()

@@

Figure \@ref(fig:death-plot) shows that the effect is more "rally 'round the flag" than "reward and punishment". That said, our model suggests at least some preliminary evidence of a transition from one mechanism to the other. In March 2020, when the pandemic started to gain pace in Britain, there was a strong positive relationship between change in the number of deaths attributable to COVID-19 and change in support for the governing Conservative Party. For every 100 extra COVID-related deaths over the past 4 weeks, Conservative support over the same period increased on average by `r britpol:::in_text((pluck(posterior_samples(death_mod, "b_deaths"), 1) + pluck(posterior_samples(death_mod, "b_time:deaths"), 1)*2)*100, inside = F, text = " points")`. Six months later, in September 2020, this relationship had weakened markedly. Though still positive, it had fallen to only `r britpol:::in_text((pluck(posterior_samples(death_mod, "b_deaths"), 1) + pluck(posterior_samples(death_mod, "b_time:deaths"), 1)*8)*100, inside = F, text = " points")`. A year after the pandemic began, the relationship had reversed. In March 2021, every 100 additional COVID-related deaths over the past 4 weeks led Conservative support to fall on average by `r britpol:::in_text((pluck(posterior_samples(death_mod, "b_deaths"), 1) + pluck(posterior_samples(death_mod, "b_time:deaths"), 1)*14)*100, inside = F, text = " points")`.

@@

cond_fx %>% 
  ggplot(
    aes(
      x = deaths,
      y = estimate__,
      ymin = estimate__ - qnorm(.975)*se__,
      ymax = estimate__ + qnorm(.975)*se__
    )
  ) +
  facet_wrap(~ cond__) +
  geom_vline(
    xintercept = 0,
    colour = britpol:::bailey_colours("grey4"),
    linetype = "dotted"
  ) +
  geom_ribbon(
    colour = NA,
    fill = britpol:::bailey_colours("red"),
    alpha = .2) +
  geom_line(colour = britpol:::bailey_colours("red")) +
  scale_y_continuous(
    breaks = seq(-.1, .1, by = .05),
    labels = scales::percent_format(suffix = "pts", accuracy = 1)
  ) +
  scale_x_continuous(
    breaks = seq(-10, 10, by = 5),
    labels = scales::number_format(scale = 100)
  ) +
  coord_cartesian(
    ylim = c(-.1, .1),
    xlim = c(-10, 10)
  ) +
  labs(
    y = "Change in Con. Support, Past 4 Weeks",
    x = "Change in Deaths Attributable to COVID-19, Past 4 Weeks"
  ) +
  theme_bailey()
```

@@

britpol:::save_info(path = here("sessions", "007_paper.txt"))
